---
title: "SRout Assign 15"
author: "Subhalaxmi Rout"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty: 
    theme: architect
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<span style="color:darkred">
1. Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary. <br>
( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )
</span>

<span style="color:darkblue">
**Solution**
</span>

```{r}
x <- c(5.6,6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)
data <- as.data.frame(cbind(x,y))
lm <- lm(y~x, data = data)
summary(lm)
plot(data$x,data$y, xlab = "x", ylab = "y")
abline(lm, col = 'darkred')
```

Regression line equation y = 4.257*x - 14.8

<span style="color:darkred">
2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form
( x, y, z ). Separate multiple points with a comma. <br>
$$f ( x, y ) = 24x - 6xy^2 - 8y^3$$
</span>


<span style="color:darkblue">
**Solution**
</span>

$$z = 24x - 6xy^2 - 8y^3$$
$$\frac {dz}{dx} = 24 - 6y^2$$
$$\frac {dz}{dy} = - 12xy - 24y^2$$

$$\frac {dz}{dydx} = - 12y$$


$$ 24 - 6y^2 = 0, \space 6y^2 = 24, \space y^2 = 4, \space y = (+2,-2)$$

$$- 12xy - 24y^2 = 0, \space -xy -2y^2 = 0, \space x = -2y, \space x = (+4,-4)$$
We have critical point (4, -2) and (-4, 2).

D = Second derivative test

* if D(x,y) > 0 and f(x,y) > 0 then f(x,y) is local minimum
* if D(x,y) > 0 and f(x,y) < 0 then f(x,y) is local maximum
* if D(x,y) < 0 then f(x,y) is saddle point

$$ \frac {d^2z}{d^2x} = 24 - 6y^2 = 0$$

$$ \frac {d^2z}{d^2y} = - 12xy - 24y^2$$
$$ \frac {d^2z}{d^2y} = - 12x - 48y$$


$$D(x,y) = fxx(x,y)fyy(x,y)−fxy(x,y)^2 = 0 * (- 12x - 48y) - (-12y)^2 = -144y^2 (which \space is  \space < 0)$$
$$for (4, -2) = -144 *(-2^2) = -144 * 4 = - 576 < 0$$

$$for (-4, 2) = -144 *(2^2) = -144 * 4 = - 576 < 0$$
So, saddle points: (4, -2) and (-4, 2).


<span style="color:darkred">
3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she
sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the
"house" brand and 40 + 11x - 23y units of the "name" brand. <br>
Step 1. Find the revenue function R ( x, y ). <br>
Step 2. What is the revenue if she sells the "house" brand for 2.30 dollar and the "name" brand for $4.10?
</span>

<span style="color:darkblue">
**Solution**
</span>

Step1:

Revenue(x,y) = house brand sells* dollar +  name brand sell * dollar

$$R(x,y) = x*(81 - 21x + 17y) + y*(40 + 11x - 23y)$$
Step 2:

```{r}
x <- 2.30
y <- 4.10
revenue <- function(x,y){
  return(x*(81 - 21*x + 17*y) + y*(40 + 11*x - 23*y))
}
revenue(x,y)
```
<span style="color:darkred">
4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by $$C(x, y) = \frac {1}{6} x^2 +  \frac{1}{6}y^2 + 7x + 25y + 700$$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in
each plant to minimize the total weekly cost?
</span>

<span style="color:darkblue">
**Solution**
</span>

given, x = y = 96
so, y = 96 - x
$$C(x, y) = \frac {1}{6} x^2 +  \frac{1}{6}y^2 + 7x + 25y + 700$$
$$C(x, 96-x) = \frac {1}{6} x^2 +  \frac{1}{6}(96-x)^2 + 7*x + 25*(96-x) + 700$$
Minima critical point : C′(x,96-x)=0


```{r}
eqn <- expression(1/6 *(x^2) + 1/6*(96-x)^2 + 7*x + 25*(96-x) + 700)
D(eqn, 'x')
```
```{r}
f <- function(x)  (1/6 * (2 * x) - 1/6 * (2 * (96 - x)) + 7 - 25)
uniroot(f, lower=0.1, upper=100000000)$root
```

x = 75, y = 96 -75 = 21


<span style="color:darkred">
5. Evaluate the double integral on the given region.
$$\int_R \int e^{(8x + 3y)}dA; R: 2<=x<=4 \space and \space 2<=y<=4$$
Write your answer in exact form without decimals.
</span>

<span style="color:darkblue">
**Solution**
</span>

$$\int_R \int e^{(8x + 3y)}dA; R: 2<=x<=4 \space and \space 2<=y<=4$$
$$\int_{2}^{4} \int_{2}^{4} e^{(8x + 3y)} dxdy = \int_{2}^{4} e^{(3y)} \int_{2}^{4} e^{(8x)}dxdy$$

$$ = \int_{2}^{4} e^{(3y)} \frac {1}{8}(e^{32} -e^{16})dy = \frac {1}{8} (e^{32} -e^{16}) \int_{2}^{4} e^{(3y)} dy$$

$$ = \frac {1}{8} (e^{32} -e^{16}) \frac {1}{3} e^{3y}|_{4}^{2} = \frac {1}{8} (e^{32} -e^{16}) \frac {1}{3} (e^{12}-e^{6})$$

$$ = \frac {1}{24} (e^{32} - e^{16})(e^{12} - e^{6})$$

```{r}
format(1/24*(exp(32) - exp(16)) * (exp(12) - exp(6)), scientific = F)
```