---
title: "SRout Assign 12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

### Data Desciption

The attached who.csv dataset contains real-world data from 2008. The variables included follow.

* Country: name of the country
* LifeExp: average life expectancy for the country in years
* InfantSurvival: proportion of those surviving to one year or more
* Under5Survival: proportion of those surviving to five years or more
* TBFree: proportion of the population without TB.
* PropMD: proportion of the population who are MDs
* PropRN: proportion of the population who are RNs
* PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
* GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate
* TotExp: sum of personal and government expenditures.

#### Question 1

Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, $R^2$, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

**Solution** <br>

Load the csv data in to R, this data consists of 10 coumns and 190 obsevarions.  

```{r}
who <- read.csv("/Users/subhalaxmirout/DATA 605/who.csv")
dim(who)
```
Below table shows entire data and statistical measures.

```{r}
library(DT)
DT::datatable(who)
summary(who)
```

Checking for any NA in data. 

```{r}
who[!complete.cases(who),]
```

No NA present in data set. 

```{r}
library(ggplot2)
library(scales)

ggplot(who, aes(TotExp, LifeExp)) +
  geom_point(col ="blue") +
  ylab("Avg Life Expectancy (years)") +
  xlab("Total Expenditures (Personal and Government)")  +
  theme( axis.line = element_line(colour = "darkblue", 
                      size = 1, linetype = "solid")) +
  scale_x_continuous(labels = dollar)
```

Above plot shows:

* The relationship between Life Expectancy and Total Healthcare Expenditures is not linear. 
* The high expenditures counties having high life expectancy. 

##### Create linear regression model using Expenditure and Life expectancy.

```{r}
lm = lm(LifeExp ~ TotExp,data = who)
summary(lm)
```

From Linear model we get:

* Low R-squared this means model only accounts for 25.8% of the variation in the data.
* Low p-value which is < 0.05, this means statistically significant
* Residuals showing data is not distributed linearly. 
* Standard error is 9.37

We can say, this model is not a good fit model.

#### Residual analysis

```{r}
residual <- resid(lm)
plot(who$TotExp, residual, xlab="Total Expenditures (Personal and Government)",
     ylab="Residuals",
     main="Residual Plot" ) 
abline(h = 0, col = "blue", lwd=2, lty=2)
abline(h = 10, col = "dark red", lwd=2, lty=2)
abline(h = -10, col = "dark red", lwd=2, lty=2)
```

```{r}
hist(residual, col = "sky blue")
```

```{r}
qqnorm(residual)
qqline(residual)
```

From residual analysis, we found:

* Residuals are not normally distributed and there is no constant variability.
* Histogram is unimodel and entirely left skewed
* Lots of deviations on QQ-plot

#### Question 2

Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"

**Solution** <br>

```{r}
library(dplyr)
who <- who %>%
  dplyr::mutate(TotExp_new = TotExp^.06,
         LifeExp_new = LifeExp^4.6)
```

```{r}
ggplot(who, aes(TotExp_new, LifeExp_new)) +
  geom_point(col = "blue") +
  ylab("Avg Life Expectancy (years) to the power 4.6") +
  xlab("Avg Personal and Government Expenditures (US dollars) to the power 0.06") +
  theme( axis.line = element_line(colour = "darkblue", 
                      size = 1, linetype = "solid")) +
  scale_x_continuous(labels = dollar) +
  scale_y_continuous(labels = comma)
```

#### Create another model using transformed variable

```{r}
lm_transform <- lm(LifeExp_new ~ TotExp_new,data = who)
summary(lm_transform)
```

#### Residual analysis for new model

```{r}
residual_new <- resid(lm_transform)
plot(who$TotExp_new, residual_new, xlab="Total Expenditures (Personal and Government)",
     ylab="Residuals",
     main="Residual Plot" ) 
abline(h = 0, col = "blue", lwd=2, lty=2)
```

```{r}
hist(residual_new, col = "sky blue")
```

```{r}
qqnorm(residual_new)
qqline(residual_new)
```

New transformed model, shows:

* Residuals are distributed normally and unimodal
* P-value is < 0.05, which is statistically significant
* Standard error is 90490000
* $R^2$ is 0.729, means model only accounts for 25.8% of the variation in the data
* QQ plot shows most observations are fall on the line with minimal deviation.

Out of 2 models, transformed model is better than the first model.


#### Question 3

Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.

**Solution** <br>

Using transformed model linear regression equation:
$$LifeExp ^ {4.6} = −736527910 + 620060216 ∗ TotExp ^ {0.06}$$

```{r}
forecast_life_expectancy <- function(total_exp)
  {
  result <- (-736527910 + 620060216 *(total_exp) ) ^ (1/4.6)
  return(result)
}
```

```{r}
forecast_life_expectancy(1.5)
forecast_life_expectancy(2.5)
```

The prediction at 1.5 is 63 years and 2.5 is 87 years.


#### Question 4

Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
$$LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp$$

**Solution** <br>

```{r}
lm3 <- lm(LifeExp ~ PropMD + TotExp + PropMD * TotExp, data = who)
summary(lm3)
```

From above model3, we found:

*  $R^2$ is 0.3471, means only 34% of the variance in the response variable can be explained by the independent variable.
* p-value is very small which is statisticaly significant to the model
* Standard error is 8.76

This model is not as good as the transformed model.

#### Question 5

Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

**Solution** <br>

coefficients of model 3 <br>

```{r}
summary(lm3)['coefficients']
```

```{r}
forecast_life_expectancy_2 <- function(PropMD, total_exp)
  {
  result <- 62.77270 + 1497.494 * PropMD + 0.00007233 * total_exp + 0.006025686 * PropMD * total_exp
  return(result)
}
```

```{r}
forecast_life_expectancy_2(0.03, 14)
```

The predicted value appears too high. This does not seem realistic. 