---
title: "SRout Assign3"
author: "Subhalaxmi Rout"
date: "9/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

**Problem set 1**

$(1)$ What is the rank of the matrix A?

$$A  =  \left[\begin{array}
{rrr}
1  &  2 &  3 &  4 \\
-1 &  0 &  1 &  3 \\
0  &  1 & -2 &  1 \\
5  & 4  & -2 & -3
\end{array}\right]
$$
**Solution**
```{r}
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3),nrow=4,byrow=TRUE)
det(A)
```
Matrix is invertible, so rank of the matrix is its dimention i.e 4.<br>
Using R
```{r}
#install.packages("matrixcalc")
library(matrixcalc)
cat("Rank of Matrix is ", matrix.rank(A))
```

$(2)$ Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?

**Solution** <br>
The rank of a matrix is equal to the minimum of the number of rows and the number of columns (m > n). So in this case, the maximum rank = n. <br>
Assuming matrix is non-zero, it means there is at least one element that is non-zero, Therefore the minimum rank can be 1.

$(3)$ What is the rank of matrix B?
$$B  =  \left[\begin{array}
{rrr}
1 &  2 & 1 \\
3 &  6 & 3  \\
2 &  4 & 2  
\end{array}\right]
$$
**Solution**<br>
Method - 1<br>
Using R
```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), byrow = TRUE, ncol = 3)
cat("Rank of Matrix is ", matrix.rank(B))
```
Method - 2<br>
Above matrix we can see, R3 is 2R1 and R2 is 3R1.<br>

step 1: $(R2 = R2 - 3R1)$
$$B  =  \left[\begin{array}
{rrr}
1 &  2 & 1 \\
0 &  0 & 0  \\
2 &  4 & 2  
\end{array}\right]
$$
step 2: $(R3 = R3 -2R1)$
$$B  =  \left[\begin{array}
{rrr}
1 &  2 & 1 \\
0 &  0 & 0  \\
0 &  0 & 0  
\end{array}\right]
$$
Above matrix there is only one non-zero row, so the rank of B is 1.

**Problem set 2**<br>
Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your
work. You'll need to write out the characteristic polynomial and show your solution.
$$A  =  \left[\begin{array}
{rrr}
1 &  2 & 3 \\
0 &  4 & 5  \\
0 &  0 & 6  
\end{array}\right]
$$
**Solution**<br>
To find polynomial equation, $det(A - {\lambda} I) = 0$. <br>
Where I = Identity matrix
$$I  =  \left[\begin{array}
{rrr}
1 &  0 & 0 \\
0 &  1 & 0  \\
0 &  0 & 1  
\end{array}\right]
$$


$$det(A - {\lambda} I) = \left[\begin{array}
{rrr}
1 - \lambda &  2 & 3 \\
0 &  4 - \lambda & 5  \\
0 &  0 & 6 - \lambda  
\end{array}\right] = 0
$$
Expanding along Row 3<br>
$(6 - \lambda)*(1 - \lambda)*(4 -\lambda) = 0$ <br>
$-\lambda ^ 3  + 11 \lambda ^ 2 - 34 \lambda + 24 = 0$ <br>
$\lambda ^ 3  - 11 \lambda ^ 2 + 34 \lambda - 24 = 0$ <br>
The charecterstic polynomial euqtion is,<br>
$p(A) = \lambda ^ 3  - 11 \lambda ^ 2 + 34 \lambda - 24$ <br>

From quation we get, $(6 - \lambda)*(1 - \lambda)*(4 -\lambda) = 0$ <br>
This means $\lambda = 1, 4, 6$ <br>
so eigen values are 1, 4, and 6.<br>

Calculation of eigen vectors, $\lambda = 1$ <br>

$$\left[\begin{array}
{rrr}
0 &  2 & 3 \\
0 &  3 & 5 \\
0 &  0 & 5 
\end{array}\right]
\
\left[\begin{array}
{rrr}
v1 \\
v2 \\
v3
\end{array}\right] = 0
$$
5v3 = 0<br>
v3 = 0<br>
3v2 + 5v3 = 0<br>
3v2 = 0<br>
v2 = 0<br>
For λ = 1, its eigenvectors are =<br> $$\left[\begin{array}
{rrr}
1 \\
0 \\
0
\end{array}\right]$$
For λ = 4<br>
$$\left[\begin{array}
{rrr}
-3 &  2 & 3 \\
0 &  0 & 5 \\
0 &  0 & 2 
\end{array}\right]
\
\left[\begin{array}
{rrr}
v1 \\
v2 \\
v3
\end{array}\right] = 0
$$
2v3 = 0<br>
v3 = 0<br>
-3v1 + 2v2 + 3v3 = 0<br>
-3v1 + 2v2 = 0<br>
3v1 = 2v2<br>
v1 = 2/3v2<br>
let v2 = 1 <br>
v1 = 2/3 <br>
For λ = 4, its eigenvectors are =<br>
$$\left[\begin{array}
{rrr}
2/3 \\
1 \\
0
\end{array}\right]$$
<br>
For λ = 6<br>
$$\left[\begin{array}
{rrr}
-5 &  2 & 3 \\
0 &  -2 & 5 \\
0 &  0 & 0 
\end{array}\right]
\
\left[\begin{array}
{rrr}
v1 \\
v2 \\
v3
\end{array}\right] = 0
$$
<br>
-2v2 + 5v3 = 0 <br>
2v2 = 5v3 <br>
v2 = 5/2 v3 <br>
-5v1 + 2v2 + 3v3 = 0 <br>
5v1 = 2v2 + 3v3 <br>
5v1 = 2(5/2)v3 + 3v3 <br>
5v1 = (5 + 3)v3 <br>
5v1 = 8v3 <br>
v1 = 8/5 v3 <br>
let v3 = 1 <br>
v2 = 5/2 <br>
v1 = 8/5 <br>

For λ = 6, its eigenvectors are =<br>
$$\left[\begin{array}
{rrr}
8/5 \\
5/2 \\
1
\end{array}\right]$$
<br>
Eigenvectors for matrix A are <br>
$$\left[\begin{array}
{rrr}
8/5 \\
5/2 \\
1
\end{array}\right]
\ ,
\left[\begin{array}
{rrr}
2/3 \\
1 \\
0
\end{array}\right]
\ and
\left[\begin{array}
{rrr}
1\\
0 \\
0
\end{array}\right]
$$
<br>
To cross check answer, calculate eigen vectors and eigen valus using R.
```{r}
library(pracma)
A <- matrix(c(1,2,3,0,4,5,0,0,6), byrow = TRUE, ncol = 3)
chac_poly_A <- charpoly(A, info = FALSE)
chac_poly_A
```
The charecterstic polynomial euqtion is,<br>
$p(A) = \lambda ^ 3  - 11 \lambda ^ 2 + 34 \lambda - 24$ <br>

```{r}
eig_values_A <- eigen(A)$values
cat("Eigen values: ",eig_values_A)
```
```{r}
eig_vectors_A <- eigen(A)$vectors
print("Eigen vectors: ")
eig_vectors_A
```