---
title: "SRout Assign 10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6. <br>

Find the probability that he wins 8 dollars before losing all of his money if <br>

(a) he bets 1 dollar each time (timid strategy). <br>
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy). <br>
(c) Which strategy gives Smith the better chance of getting out of jail? <br>

**Answer** <br>

(a) he bets 1 dollar each time (timid strategy) <br>

$$ P = \frac {1 - (\frac {q}{p}) ^ s } {1 - (\frac {q}{p}) ^ M} $$

Given, q = 0.6 <br>
p = 0.4, <br>
s = 1, <br>
M = 8 <br>

According to formula of Gambler’s Ruin: <br>
$$ P = \frac {1 - (\frac {q}{p}) ^ s } {1 - (\frac {q}{p}) ^ M} = \frac {1 - (\frac {0.6}{0.4}) ^ 1 } {1 - (\frac {0.6}{0.4}) ^ 8} = 0.02$$

```{r}
# using R
q <- 0.6
p <- 0.4
s <- 1
M <- 8
P <- (1 -  (q/p) ^ s) / (1 - (q/p) ^ M)
print(P)
```
The probability of wining is 0.02. <br>

(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy). <br>

If he bets his entire money each time until 8 dollar. He starts with 1 dollar and must win 3 bets in a row with probability of 0.4 <br>

At first bet, If he wins, he ends up with $2. <br>

At second bet, If he wins, he ends up with $4. <br>

At third bet, If he wins, he ends up with $8. <br>

```{r}
library(stats)
dbinom(3,3,0.4)
```

The probability of wining is 0.064. <br>

(c) The bold strategy is having higher probability than timid strategy, So, bold strategy has a better chance of getting out of jail. 