---
title: "SRout Assign 2"
author: "Subhalaxmi Rout"
date: "9/5/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem set 1

$(1)$ Show that $A^T A ≠ A A^T$ in general. (Proof and demonstration.)

**Answer** <br>
Method 1: <br>
Lets prove mathematically, <br>

$$\mathbf{A} = \left[\begin{array}
{rrr}
a & b \\
c & d 
\end{array}\right]
$$
Transpose of A
$$\mathbf{A^T} = \left[\begin{array}
{rrr}
a & c \\
b & d 
\end{array}\right]
$$
$$\mathbf{A A^T} = \left[\begin{array}
{rrr}
a & b \\
c & d 
\end{array}\right]
\
\left[\begin{array}
{rrr}
a & c \\
b & d 
\end{array}\right]
$$
$$\mathbf{A A^T}= \left[\begin{array}
{rrr}
(a^2 + b^2) & (ac + bd) \\
(ac + bd) & (c^2 + d^2) 
\end{array}\right]
$$


Lets find out $A^T A$

$$\mathbf{A^T A} =\left[\begin{array}
{rrr}
a & c \\
b & d 
\end{array}\right]
\
\left[\begin{array}
{rrr}
a & b \\
c & d 
\end{array}\right]
$$
$$\mathbf{A^T A} =\left[\begin{array}
{rrr}
(a^2 + c^2) & (ab + cd) \\
(ab + cd) & (b^2 + d^2) 
\end{array}\right]
$$
From the above equation we can see $A^T A ≠ A A^T$

For example: 
$$\mathbf{M} = \left[\begin{array}
{rrr}
3 & 4 \\
5 & 6 
\end{array}\right]
$$
Transpose of M
$$\mathbf{M^T} = \left[\begin{array}
{rrr}
3 & 5 \\
4 & 6 
\end{array}\right]
$$

$$M M^T =\left[\begin{array}
{rrr}
3 & 4 \\
5 & 6 \\
\end{array}\right]
\
\left[\begin{array}
{rrr}
3 & 5 \\
4 & 6 \\
\end{array}\right]
$$


$$M M^T  =  \left[\begin{array}
{rrr}
(3^2 + 4^2) & (3*5 + 4*6) \\
(3*5 + 4*6) & (5^2 + 6^2) 
\end{array}\right]
$$




$$M M^T  =  \left[\begin{array}
{rrr}
(9 + 16) & (15 + 24) \\
(15 + 24) & (25 + 36) 
\end{array}\right]
$$


$$M M^T  =  \left[\begin{array}
{rrr}
25 & 39 \\
39 & 61 
\end{array}\right]
$$

Similary product of $M^T M$
$$M^T M = \left[\begin{array}
{rrr}
3 & 5 \\
4 & 6 \\
\end{array}\right]
\
\left[\begin{array}
{rrr}
3 & 4 \\
5 & 6 \\
\end{array}\right]
$$

$$M^T M  =  \left[\begin{array}
{rrr}
(3^2 + 5^2) & (3*4 + 5*6) \\
(3*4 + 5*6) & (4^2 + 6^2) 
\end{array}\right]
$$


$$M^T M  =  \left[\begin{array}
{rrr}
(9 + 25) & (12 + 30) \\
(12 + 30) & (16 + 36) 
\end{array}\right]
$$

$$M^T M  =  \left[\begin{array}
{rrr}
(9 + 25) & (12 + 30) \\
(12 + 30) & (16 + 36) 
\end{array}\right]
$$


$$M^T M  =  \left[\begin{array}
{rrr}
34 & 42 \\
42 & 52 
\end{array}\right]
$$

Above example we see $A^T A ≠ A A^T$.

Method 2:  Using R<br>
```{r}
A = matrix(c(3,5,4,6), nrow = 2, ncol = 2)
TA = t(A)
TA
```
$A A^T$
```{r}
AAT <- A %*% TA
AAT
```
$A^T A$
```{r}
ATA <- TA %*% A
ATA
```
Hence, from above example we see $A^T A ≠ A A^T$.

$(2)$ For a special type of square matrix A, we get $A^T A = A A^T$ . Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

**Answer** <br>
Method 1 :<br>
Suppose $A = A^T$ <br>
For example:<br>
$$A  =  \left[\begin{array}
{rrr}
1 & 3 \\
3 & 1 
\end{array}\right]
$$
Transpose of A :<br>
$$A^T  =  \left[\begin{array}
{rrr}
1 & 3 \\
3 & 1 
\end{array}\right]
$$
$$A A^T = \left[\begin{array}
{rrr}
1 & 3 \\
3 & 1 \\
\end{array}\right]
\
\left[\begin{array}
{rrr}
1 & 3 \\
3 & 1 \\
\end{array}\right]
$$
$$A A^T  = \left[\begin{array}
{rrr}
(1*1 + 3*3) & (1*3 + 3*1) \\
(3*1 + 1*3) & (3*3 + 1*1) \\
\end{array}\right]
$$
$$A A^T  = \left[\begin{array}
{rrr}
10 & 6 \\
6 & 10 \\
\end{array}\right]
$$
Similary $A^T A$

$$A^T A = \left[\begin{array}
{rrr}
1 & 3 \\
3 & 1 \\
\end{array}\right]
\
\left[\begin{array}
{rrr}
1 & 3 \\
3 & 1 \\
\end{array}\right]
$$


$$A^T A  = \left[\begin{array}
{rrr}
(1*1 + 3*3) & (1*3 + 3*1) \\
(3*1 + 1*3) & (3*3 + 1*1) \\
\end{array}\right]
$$

$$A^T A  = \left[\begin{array}
{rrr}
10 & 6 \\
6 & 10 \\
\end{array}\right]
$$

So, we can confirm that when $A = A^T$ then $A A^T = A^T A$.<br>

Method 2: <br>
Lets varify the same using R.<br>

```{r}
A <-  matrix(c(1,3,3,1), ncol=2, nrow = 2) 
T_A = t(A)
T_A
```
$A A^T$
```{r}
AT_A <- A %*% T_A
AT_A
```
$A^T A$
```{r}
T_AA <- T_A %*% A
T_AA
```
$A A^T$ and $A^T A$ gives the same result. <br>

**Problem set 2** <br>
Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track 
ights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your 
ight using radars.<br>
Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer. Please submit your response in an R Markdown document using our class naming
convention, E.g. LFulton_Assignment2_PS2.png<br>
You don't have to worry about permuting rows of A and you can assume that A is less
than 5x5, if you need to hard-code any variables in your code. If you doing the entire
assignment in R, then please submit only one markdown document for both the problems.

```{r}
factorize <- function(A) {
  
  U <- A
  n <- nrow(A)
  L <- diag(n)
  
  # If 1x1 matrix, then U = A and L=[1]
  if (n == 1) 
    {
    return(list(L,U))
    }
  
  
  # Determine multiplier for each position and add it to L
  for(k in 1:(n-1)) 
  {
    for (i in (k+1):n) { 
      L[i,k] <- U[i,k]/U[k,k]
      U[i,] <- U[i,] - L[i,k] * U[k,]
    }
  }
  return(list(L,U))
}

# Test with 3x3 matrix
Mat <- matrix(c(1,5,8,-7,6,2,9,3,4), nrow=3, byrow=TRUE)
LU <- factorize(Mat)
cat("Lower trangualr matrix :", "\n")
LU[[1]]
cat("Upper trangular matrix", "\n")
LU[[2]]

# To verify the answer 
Mat == LU[[1]] %*% LU[[2]]
```